---
title: "Final Project- Repeatable Experiments"
author: "Rebecca Kotula"
date: "8/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load Packages
```{r, warning=FALSE}
source("cross_validation_functions.R")# custom CV & data-loading functions
library(xpectr) # set_test_seed()
library(ggplot2)
library(reshape2)
library(scales) # comma format for plots
```

## Load & Prepare Data
```{r}
# Load Data
housing <- load_housing_dataset()

# Set seed for repeatability
set_test_seed(1) 

# Split data 80/20 for training/validation, and testing
partitions <- partition(housing, p = 0.2, id_col = "id")
test.set <- partitions[[1]]
train.set <- partitions[[2]]

max.price <- max(train.set$price)
min.price <- min(train.set$price)
```

## Visualizing the Distribution of Price

```{r}
h <- ggplot(train.set, aes(x=price)) +
    geom_histogram(bins=50) +
    scale_x_continuous(labels = comma) + 
    labs(title = "Histogram of Housing Prices")
plot(h)
#ggsave("price_histogram.png", h, "png", "/Users/rebeccakotula/Library/Mobile Documents/com~apple~CloudDocs/Documents/Theory of Statistics I/Project/Images/")
```

## Correlation of Features
```{r}
correlations <- cor(train.set[,!names(train.set) %in% c("id", "df.index", "lat", "long", "date", "zipcode", "sqft_living15", "sqft_lot15")])
correlations <- round(correlations,2)
melted_cor <- melt(correlations)
c <- ggplot(data = melted_cor, aes(x=Var1, y=Var2, fill=value)) + 
    geom_tile()+
    scale_fill_gradient2(low = "red", high = "red", mid = "yellow", 
                         midpoint = 0, limit = c(-1,1), space = "Lab", 
                         name="Pearson\nCorrelation")+
    theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),
          axis.text.y = element_text(size=12),
          axis.title.x = element_blank(),
          axis.title.y = element_blank())+
    geom_text(aes(Var2, Var1, label = value), color = "black", size = 3) 
plot(c)
#ggsave("correlations.png", c, "png", "/Users/rebeccakotula/Library/Mobile Documents/com~apple~CloudDocs/Documents/Theory of Statistics I/Project/Images/", width=7,height=5)
```
```{r }
relevant.features <- sort(correlations[,1], decreasing=TRUE)
relevant.features
```

## Choice of k Comparison

Comparing multiple cross-validation methods and choices of $k$ with one static model, 100 simulations for each choice of $k$.  
```{r}
model.baseline <- "price ~ bedrooms + grade + sqft_living"

#### 10 fold cv
score.list10 <- data.frame(matrix(ncol=3,nrow=0, dimnames=list(NULL, c("KCV.RMSE", "MPCV.RMSE", "LOOCV.RMSE"))))

for (i in 1:100){
    # Set seed for repeatability
    set_test_seed(i) 
    # Create fold assignments
    train.set10 <- fold(train.set, 10, id_col="id")
    # Scores with 10 folds
    k.score10 <- KCV(train.set10, k=10, model=model.baseline, predict.col="price")
    m.score10 <- MPCV(train.set10, k=10, model=model.baseline, predict.col="price", df.index = "df.index")
    l.score10 <- LOOCV(train.set10, model=model.baseline, predict.col="price")
    
    score.list10 <- rbind(score.list10, data.frame(KCV.RMSE=k.score10, MPCV.RMSE=m.score10, LOOCV.RMSE=l.score10))
}

print("10 Fold scores")
print(colMeans(score.list10))


#### 5 fold cv
score.list5 <- data.frame(matrix(ncol=3,nrow=0, dimnames=list(NULL, c("KCV.RMSE", "MPCV.RMSE", "LOOCV.RMSE"))))

for (i in 1:100){
    # Set seed for repeatability
    set_test_seed(i) 
    # Create fold assignments
    train.set5 <- fold(train.set, 5, id_col="id")
    # Scores with 5 folds
    k.score5 <- KCV(train.set5, k=5, model=model.baseline, predict.col="price")
    m.score5 <- MPCV(train.set5, k=5, model=model.baseline, predict.col="price", df.index = "df.index")
    l.score5 <- LOOCV(train.set5, model=model.baseline, predict.col="price")
    
    score.list5 <- rbind(score.list5, data.frame(KCV.RMSE=k.score5, MPCV.RMSE=m.score5, LOOCV.RMSE=l.score5))
}
print("5 Fold Scores")
print(colMeans(score.list5))
```

## Stepwise Feature Elimination to Compare CV Methods

Set up the dataset, using 5 folds. 
```{r}
# Set seed for repeatability
set_test_seed(1) 
# Create fold id column in train set (used by custom CV functions)
train.set1 <- fold(train.set, 5, id_col="id")
```

### Uninformed Baseline (features not sorted)
```{r, warning=FALSE}
###### Models ##################################################################
#unused cols: "sqft_living15", "sqft_lot15", "zipcode"
model1 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors + waterfront + view + condition + grade + sqft_above + sqft_basement + yr_built + yr_renovated"
model2 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors + waterfront + view + condition + grade + sqft_above + sqft_basement + yr_built"
model3 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors + waterfront + view + condition + grade + sqft_above + sqft_basement"
model4 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors + waterfront + view + condition + grade + sqft_above"
model5 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors + waterfront + view + condition + grade"
model6 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors + waterfront + view + condition"
model7 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors + waterfront + view "
model8 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors + waterfront"
model9 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot + floors"
model10 <- "price ~ bedrooms + bathrooms + sqft_living+ sqft_lot"
model11 <- "price ~ bedrooms + bathrooms + sqft_living"
model12 <- "price ~ bedrooms + bathrooms"
model13 <- "price ~ bedrooms"
model_list <- c(model1, model2, model3, model4, model5, model6, model7, model8, model9, model10, model11, model12, model13)


###### Run CV ##################################################################
k.scores <- c()
m.scores <- c()
l.scores <- c()
for (model in model_list) {
    k.scores <- c(k.scores, KCV(train.set1, k=10, model=model, predict.col="price"))
    m.scores <- c(m.scores, MPCV(train.set1, k=10, model=model, predict.col="price", df.index = "df.index"))
    l.scores <- c(l.scores, LOOCV(train.set1, model=model, predict.col="price"))
}

results <- data.frame(model=seq(1,length(model_list), 1), k.scores, m.scores, l.scores)
#print(which.min(k.scores))
results
saveRDS(results,file="/Users/rebeccakotula/Library/Mobile Documents/com~apple~CloudDocs/Documents/Theory of Statistics I/Project/Datasets/results.Rda")
```
Which model is has the lowest score for each of the three methods? (kcv, mpcv, loocv)
```{r}
best.k <- which.min(results$k.scores)
best.m <- which.min(results$m.scores)
best.l <- which.min(results$l.scores)

print(c(best.k, best.m, best.l))
```

Plot of the results:
```{r}
results.melt <- melt(results, id.var=c("model"))
colnames(results.melt) <- c("model", "cv_method", "score")
g <- ggplot(data = results.melt, aes(x=model, y=score, col=cv_method)) +
    geom_line()+
    labs(title = "Comparison of CV Methods (unsorted)", x = "Model Label", y = "RMSE", color = "CV Method") +
    scale_color_manual(labels = c("KCV", "MPCV", "LOOCV"), values = c("darkorange1", "darkolivegreen4", "cornflowerblue")) +
    scale_x_continuous(labels=as.character(seq(1,length(model_list), 1)), breaks=seq(1,length(model_list), 1))
plot(g)
#ggsave("unsorted_stepwise_comparisons.png", g, "png", "/Users/rebeccakotula/Library/Mobile Documents/com~apple~CloudDocs/Documents/Theory of Statistics I/Project/Images/")
```

### Sorted Baseline (by correlation)
```{r, warning=FALSE}
###### Models ##################################################################
#unused cols: "sqft_living15", "sqft_lot15", "zipcode"
model1 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view + sqft_basement + bedrooms + waterfront + floors + yr_renovated + sqft_lot + yr_built + condition"
model2 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view + sqft_basement + bedrooms + waterfront + floors + yr_renovated + sqft_lot + yr_built"
model3 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view + sqft_basement + bedrooms + waterfront + floors + yr_renovated + sqft_lot"
model4 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view + sqft_basement + bedrooms + waterfront + floors + yr_renovated"
model5 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view + sqft_basement + bedrooms + waterfront + floors"
model6 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view + sqft_basement + bedrooms + waterfront"
model7 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view + sqft_basement + bedrooms"
model8 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view + sqft_basement"
model9 <- "price ~ sqft_living+ grade + sqft_above + bathrooms + view"
model10 <- "price ~ sqft_living+ grade + sqft_above + bathrooms"
model11 <- "price ~ sqft_living+ grade + sqft_above"
model12 <- "price ~ sqft_living+ grade"
model13 <- "price ~ sqft_living"

model_list <- c(model1, model2, model3, model4, model5, model6, model7, model8, model9, model10, model11, model12, model13)


###### Run CV ##################################################################
k.scores <- c()
m.scores <- c()
l.scores <- c()
for (model in model_list) {
    k.scores <- c(k.scores, KCV(train.set1, k=10, model=model, predict.col="price"))
    m.scores <- c(m.scores, MPCV(train.set1, k=10, model=model, predict.col="price", df.index = "df.index"))
    l.scores <- c(l.scores, LOOCV(train.set1, model=model, predict.col="price"))
}

results2 <- data.frame(model=seq(1,length(model_list), 1), k.scores, m.scores, l.scores)
#print(which.min(k.scores))
results2
saveRDS(results2,file="/Users/rebeccakotula/Library/Mobile Documents/com~apple~CloudDocs/Documents/Theory of Statistics I/Project/Datasets/results_sorted.Rda")

```
Which model is has the lowest score for each of the three methods? (kcv, mpcv, loocv)
```{r}
best.k <- which.min(results2$k.scores)
best.m <- which.min(results2$m.scores)
best.l <- which.min(results2$l.scores)

print(c(best.k, best.m, best.l))
```

Plot of the results:
```{r}
results2.melt <- melt(results2, id.var=c("model"))
colnames(results2.melt) <- c("model", "cv_method", "score")
g <- ggplot(data = results2.melt, aes(x=model, y=score, col=cv_method)) +
    geom_line()+
    labs(title = "Comparison of CV Methods (sorted)", x = "Model Label", y = "RMSE", color = "CV Method") +
    scale_color_manual(labels = c("KCV", "MPCV", "LOOCV"), values = c("darkorange1", "darkolivegreen4", "cornflowerblue")) +
    scale_x_continuous(labels=as.character(seq(1,length(model_list), 1)), breaks=seq(1,length(model_list), 1))
plot(g)
#ggsave("sorted_stepwise_comparisons.png", g, "png", "/Users/rebeccakotula/Library/Mobile Documents/com~apple~CloudDocs/Documents/Theory of Statistics I/Project/Images/")
```




